---
title: "Bronx Exploratory Analysis"
author: "Anthony Dalton"
date: "October 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading in the Bronx data and doing some cursory glances at the data

Reading in the raw data, calling the player library and running head, summary, and str.

```{r read}
setwd("C:/Users/adalton/Documents/Week6LiveSession/data/")
library(plyr)
##skipping the first four header lines
bronx = read.csv("rollingsales_bronx.csv", skip=4)
## Check the data
head(bronx)
summary(bronx)
str(bronx) # Very handy function!
```

## Data Cleanup

clean/format the data with regular expressions
pattern "[^[:digit:]]" refers to members of the variable name that
start with digits. We use the gsub command to replace them with a blank space.
we create a new variable that is a "clean' version of sale.price.
And sale.price.n is numeric, not a factor.
```{r cleanup}
bronx$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", bronx$SALE.PRICE))
count(is.na(bronx$SALE.PRICE.N))
names(bronx) <- tolower(names(bronx)) # make all variable names lower case
## Get rid of leading digits
bronx$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", bronx$gross.square.feet))
bronx$land.sqft <- as.numeric(gsub("[^[:digit:]]","", bronx$land.square.feet))
bronx$year.built <- as.numeric(as.character(bronx$year.built))
```

## Continued data cleanup, summary of valid sale prices, and histogram on sale prices
```{r}
bronx.sale <- bronx[bronx$sale.price.n!=0,]
##plot(bronx.sale$gross.sqft,bronx.sale$sale.price.n)
##plot(log10(bronx.sale$gross.sqft),log10(bronx.sale$sale.price.n))
bronx.sale <- bronx[bronx$sale.price.n!=0,]
##plot(bronx.sale$gross.sqft,bronx.sale$sale.price.n)
##plot(log10(bronx.sale$gross.sqft),log10(bronx.sale$sale.price.n))
## for now, let's look at 1-, 2-, and 3-family homes
bronx.homes <- bronx.sale[which(grepl("FAMILY",bronx.sale$building.class.category)),]
dim(bronx.homes)
##plot(log10(bronx.homes$gross.sqft),log10(bronx.homes$sale.price.n))
summary(bronx.homes[which(bronx.homes$sale.price.n<100000),])
## remove outliers that seem like they weren't actual sales
bronx.homes$outliers <- (log10(bronx.homes$sale.price.n) <=5) + 0
bronx.homes <- bronx.homes[which(bronx.homes$outliers==0),]
hist(bronx.homes$sale.price.n)
##plot(log10(bronx.homes$gross.sqft),log10(bronx.homes$sale.price.n))
```

## Facts on sale prices by neighborhood
Mean and Standard Deviation using clean dataset
```{r}
ddply(bronx.homes,~neighborhood,summarise,mean=mean(sale.price.n),sd=sd(sale.price.n))
```

## Scatterplot on square footage and sale price
There is a correlation between a larger home and sale price.
```{r}
summary(bronx$gross.sqft)
plot(bronx.homes$gross.sqft, bronx.homes$sale.price.n, main="Correlation between square footage and sale price")

```
